---
title: "Orders Analysis"
subtitle: <u>1. Practical examples of using R in Supply Chain Management.</u>
author: "**autor**: Maciej Lecicki"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, warning = FALSE, message = FALSE, fig.pos = "H")
```


Association Rules, commonly called Market Basket Analysis is a very helpful technique commonly used by retailers to uncover associations between items. \newline
The idea is to define certain (probability) rules and on that basis identify likelihood that given products' (items) will appear in the same transaction (basket).
Most common and intuitive example of products that are bought together is bread and butter.

Market Basket Analysis is based on three key measures: Support, Lift and Confidence (4th measure, Expected Confidence, can be also considered) and therefore is easy understand and outcome is easy to explain to stakeholders. \newline
Description of key measures: 

* Support can be desribed as popularity of item(set); it's a fraction (percentage) of all transactions that contain given item(set),
* Confidence tells us how often the rule is true; it shows the percentage in which item B is bought together with item A,
* Lift is an indicator of association strength: 
  - lift > 1: B is likely to be occur together with A
  - lift < 1: B is unlikely to occurr together with A


The key challenge when performing Market Basket was the size of data to process. Recently it has become less problematic due to increase of computational power, however it can still be a problem when summarizing or visualizing results if rules parameters are incorrect.\newline

Market Basket Analysis can also be applied to supply chain management, specifically to logistics and warehousing. \newline 
The scope of the analysis in this chapter is customers orders dispatched from 4 different consolidation DCs. The purpose is to investigate if there are any associations between products which can be foundation for SKUs allocation to consolidation DC and further grouping them together at particular warehouse location (shelf) in order to streamline distribution process and reduce warehouse handling costs related to dispatch process.\newline


Let's first upload required packages.

```{r ch4_lib_data, echo = TRUE, include = TRUE}
library(arules)
library(arulesCBA)
library(arulesViz)
library(htmlwidgets)
library(tidyverse)
```


Next step is to bring in data with customers' orders in required format.
We'll use read.transaction function from arules library.

```{r data_upload, echo = TRUE, include = TRUE}
market_order <- read.transactions("data/basket_data.csv",
                                  format = "single",
                                  sep = ",",
                                  header = TRUE,
                                  cols = c("Document", "Product"))
```


\pagebreak
We can summarize the data to display some simple statistics like number of transactions (orders), their size or most frequent items (SKUs).

```{r basket_summary, echo = TRUE, include = TRUE}
summary(market_order)
```
\pagebreak
There's 844 transactions (customer orders) and 486 items (products).\newline
We can see that there's a handful of orders with only one item. 
Let's remove them.


```{r remove_orders_with_one_item}

# remove transactions with 1 Item

market_order_filter <- market_order[size(market_order) > 1]

length(market_order_filter)
size(market_order_filter)
inspect(head(market_order_filter))
```

```{r brands_basic_statistics, include = TRUE}
summary(market_order_filter)
```


Before we perform MBA, we need to define values of parameters for association rules (in other words, we need to decide on measures values). \newline
It's an important step as it will define the number of association rules and therefore granularity of analysis and its outcome. It is recommended to use expertise knowledge to find the optimal number applicable to a given business environment.\newline

This task is usually done on the basis of Confidence and Support. Let's define a grid of parameters and investigate thresholds that will help us select optimal number of rules.\newline
Results are presented in Figure 49.

```{r parameters_grid}
# selecting support & confidence level parameters

supp_lev <- seq(from = 0.01, to = 0.2, by = 0.01)
conf_lev <- seq(from = 0.1, to = 0.9, by = 0.05)

par_check <- expand.grid(supp_level = supp_lev, 
                         conf_level = conf_lev)

par_check <- cbind(par_check,
                   count = NA)

for (i in 1:nrow(par_check)) {
  par_check[i, 3] =
    length(apriori(market_order_filter,
                   parameter = list(supp = par_check[i, 1],
                                    conf = par_check[i, 2],
                                    target = "rules")))
}


par_check_filter <- par_check %>%
  as_tibble() %>%
  filter(count > 10) %>%
  mutate(
    supp_level = as.character(supp_level),
    conf_level = as.character(conf_level)) %>%
  unite(col = par_pair,
        supp_level, 
        conf_level,
        remove = FALSE)
```

```{r threshold_plot, include = TRUE, fig.cap = "Number of association rules between Brands based on combination of Support and Confidence levels.", out.extra = ''}

ggplot(par_check_filter,
       aes(x = par_pair,
           y = count)) +
  geom_col() +
  theme(
    axis.text.x = element_text(angle = 90,
                               size = 7)
  ) +
  labs(title = "Number of association rules by parameters value",
       x =  "Support_Confidence levels",
       y = "Number of association rules")

```

<!-- market_order <- read.transactions("data/basket_data.csv", -->
<!--                                   format = "single", -->
<!--                                   sep = ",", -->
<!--                                   header = TRUE, -->
<!--                                   cols = c("Document", "Product")) -->

<!-- length(market_order) -->

<!-- size(market_order) -->

<!-- inspect(head(market_order)) -->

<!-- # remove transactions with 1 Item -->

<!-- market_order_filter <- market_order[size(market_order) > 1] -->

<!-- length(market_order_filter) -->

<!-- size(market_order_filter) -->

<!-- inspect(head(market_order_filter)) -->


<!-- summary(market_order_filter) -->

<!-- # co-occurence matrix - chiSquared: coexistance test which returns p-value -->
<!-- # HO: co-occurence of rows & columns (items) is independent (p-value > alpha) -->
<!-- # HA: co-occurence of rows & columns (items) is dependent of each other (p-value < alpha) -->

<!-- crossTable(market_order_filter, measure = "chiSquared", sort = TRUE)[1:8, 1:8] -->
<!-- # co-occurance of some items is dependent of each other -->

<!-- # let's see transactions sparsity -->
<!-- image(market_order_filter) -->

<!-- # most frequent items -->
<!-- crossTable(market_order_filter, sort = TRUE)[1:10, 1:10] -->

<!-- # frequency plots -->
<!-- itemFrequencyPlot(market_order_filter, -->
<!--                   topN = 10, -->
<!--                   main = "Absolute Product Frequency", -->
<!--                   type = "absolute", -->
<!--                   horiz = TRUE, -->
<!--                   col = "orange") -->

<!-- # barplot(sort(table(unlist(LIST(market_order_filter))), decreasing = TRUE)[1:10], -->
<!-- #         horiz = TRUE, las = 1, col = "steelblue3") -->

<!-- # Key association measures: -->

<!-- # 1. Support (popularity of item(set) ) -->
<!-- #    fraction (percentage) of all transactions that contain given item(set) -->

<!-- # 2. Confidence (how often the rule is true) -->
<!-- #    shows the percentage in which Y is bought together with X -->

<!-- # 3. Lift (how strong is the association) -->
<!-- #   lift > 1: Y is likely to be bought when X is bougth -->
<!-- #   lift < 1: Y is unlikely to be bought when X is bougth -->

<!-- # Count tells how many times given item(set) appeared in all transactions -->

<!-- # selecting support & confidence level parameters -->

<!-- supp_lev <- seq(from = 0.005, to = 0.02, by = 0.001) -->
<!-- conf_lev <- seq(from = 0.1, to = 0.9, by = 0.05) -->

<!-- par_check <- expand.grid(supp_level = supp_lev,  -->
<!--                          conf_level = conf_lev) -->

<!-- par_check <- cbind(par_check, -->
<!--                    count = NA) -->

<!-- for (i in 1:nrow(par_check)) { -->
<!--   par_check[i, 3] = -->
<!--     length(apriori(market_order_filter, -->
<!--                    parameter = list(supp = par_check[i, 1], -->
<!--                                   conf = par_check[i, 2], -->
<!--                                   target = "rules"))) -->
<!-- } -->


<!-- par_check_filter <- par_check %>% -->
<!--   as_tibble() %>% -->
<!--   filter(count > 10) %>% -->
<!--   mutate( -->
<!--     supp_level = as.character(supp_level), -->
<!--     conf_level = as.character(conf_level)) %>% -->
<!--   unite(col = par_pair, -->
<!--           supp_level,  -->
<!--           conf_level, -->
<!--           remove = FALSE) -->

<!-- ggplot(par_check_filter, -->
<!--        aes(x = par_pair, -->
<!--            y = count)) + -->
<!--   geom_col() + -->
<!--   theme( -->
<!--     axis.text.x = element_text(angle = 90, -->
<!--                                size = 7) -->
<!--   ) + -->
<!--   labs(title = "Number of association rules by parameters value", -->
<!--        x =  "Support_Confidence levels", -->
<!--        y = "Number of association rules") -->

<!-- # apriori function for frequent itemsets with -->
<!-- # support = 0.01, -->
<!-- # confidence = 0.3 -->

<!-- trans_frequent = apriori(market_order_filter, -->
<!--                          parameter = list( -->
<!--                            supp = 0.01, -->
<!--                            conf = 0.3, -->
<!--                            minlen = 2, -->
<!--                            target = "frequent itemsets") -->
<!--                          ) -->

<!-- # retrieve results -->
<!-- inspect(sort(trans_frequent, by = "support")) -->

<!-- # apriori function for rules -->

<!-- trans_rules = apriori(market_order_filter, -->
<!--                          parameter = list( -->
<!--                            supp = 0.01, -->
<!--                            conf = 0.3, -->
<!--                            minlen = 2, -->
<!--                            target = "rules"), -->
<!--                       control = list(verbose = FALSE) -->
<!-- ) -->

<!-- inspect(sort(trans_rules, by = "lift")) -->

<!-- # subsetting rules (example) -->

<!-- inspect(subset(trans_frequent, -->
<!--                subset = items %in% c("A338HC") & support > 0.01) -->
<!--         ) -->

<!-- # interactive visualization of rules -->

<!-- # summary -->
<!-- plot(trans_rules, engine = "plotly") -->

<!-- #interactive product selection -->
<!-- inspectDT(trans_rules) -->

<!-- #graph -->
<!-- plot(trans_rules, method = "graph", -->
<!--      engine = "htmlwidget") -->

<!-- # to save widget -->

<!-- # rules_html = plot(trans_rules, method = "graph", -->
<!-- #                   engine = "htmlwidget") -->
<!-- # saveWidget(rules_html, file = "trans_rules.html") -->
<!-- # saveAsGraph(rules_html, file = "trans_rules.graphml") -->

<!-- # grouped rules -->
<!-- plot(trans_rules, method = "grouped", measure = "lift", shading = "confidence") -->

<!-- # parallel coordinates plot -->
<!-- plot(trans_rules, method = "paracoord") -->
<!-- # not applicable here as association rules are 1:1 -->

<!-- # Shiny app -->
<!-- ruleExplorer(trans_rules) -->